---
title: Harvesting Real Estate
author: ''
date: '2021-05-12'
slug: rvesting-real-estate
categories: ["R", "Barbados"]
tags: ["Real Estate", "rvest", "web scraping"]
subtitle: ''
summary: ''
authors: []
lastmod: '2021-05-12 23:12:40 -04'
featured: no
image:
  caption: 'Image credit: [**Unsplash**](https://unsplash.com/photos/f9qZuKoZYoY)'
  focal_point: ''
  preview_only: no
projects: []
---
## The Intro

I find one of the strange road bumps you don't expect in practicing data analytics is finding the data in the first place. Yes the world is currently overflowing with with information, 

## The Data

The dashboard allows us to download data from their available stations. What we'll be doing is downloading their irradiance information.

Now this setup supposedly has multiple  [ways](https://earthcubeprojects-chords.github.io/chords-docs/usingchords/#retrieving-data) to download data, but I seem to be currently limited to the old fashioned download from dashboard, so that's exactly what I did.

I've manually downloaded the irradiance for each day going back a couple of days. The code will simply be made to handle either one or multiple files.

### The Code

We'll start by just loading our libraries.
```{r Load Libraries, message = FALSE}
library(tidyverse)
```


#### The Variables

These are the variables which will select the folder and files I wish to look at. In this way it should be easy to see I can easily change the code to red multiple file types, or variable with litlle hassle.

```{r Set Variables}
  file_type <- "csv"
  radiation_folder <- "radiation"
  radiation_data <- "grafana"
```


```{r Download Files}

# Get possible files
files <- list.files(full.names = TRUE,
                    recursive = TRUE,
                    include.dirs = TRUE)

# Select Desired folder
used_file_radiation <- files[str_detect(files, radiation_folder)]
# Select Desired file
used_file_radiation <- files[str_detect(files, radiation_data)]
# Keep only the files which end with the given pattern
used_file_radiation <- used_file_radiation[str_detect(string = used_file_radiation,
                                                      pattern = paste0("(",file_type,")$"))]

# Download File(s) & transform
if(length(used_file_radiation) == 1){
  
  
  radiation_raw <- read_delim(used_file_radiation,
                              col_types = cols(Time = col_character(),
                                               Value = col_double()),
                              delim = ";",
                              na = "null") 
  
} else {
  
  radiation_raw <- used_file_radiation %>%
    map_df(~read_delim(.,
                       col_types = cols(Time = col_character(),
                                        Value = col_double()),
                       delim = ";",
                       na = "null")
    )
  
} 

radiation_raw <- radiation_raw %>%
  distinct() %>%
  # Split Datetime into Date and Time
  separate(col = Time,
           into = c("Date", "Time"),
           sep = "T") %>%
  # Discard Time Zone as it's not needed
  separate(col = Time,
           into = c("Time", NA),
           sep = "-") %>%
  # Convert date and time to correct format
  mutate(Date = lubridate::ymd(Date),
         Time =  hms::as_hms(Time),
         # Add back in Date - Time column in case it's useful
         Date_Time = lubridate::ymd_hms(paste0(Date,"T",Time)),
         Month = lubridate::month(Date, label = TRUE)
  ) %>%
  select(Series, Month, Date_Time, everything()) %>%
  arrange(Date_Time)

```


#### The Plot

Just to see that we've downloaded what we wanted, we can create a quick plot.
```{r Plot Raw Data, fig.height = 10, fig.width = 10}

radiation_raw %>%
  filter(!is.na(Value)) %>%
  ggplot(aes( x = Date_Time, y = Value)) +
  geom_line(aes(colour = Series)) +
  facet_grid(Series~.,
             scales = "free_y") +
  labs(title = 'The Daily Irradiance',
       x = 'Time',
       y = 'Downwelling Irradiance (W/m^2)') +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "none")
```

Well it does get what we want, but it looks a bit chaotic. Let's instead plot the daily average for each month and see if that makes it easier to digest.

```{r Plot Summarised Data, fig.height = 10, fig.width = 10}

radiation_raw %>%
  filter(!is.na(Value)) %>%
  group_by(Month, Time, Series) %>%
  summarise(Radiation_mean = mean(Value),
            Radiation_sd = sd(Value)) %>%
  ungroup() %>%
  ggplot(aes(x = Time, y = Radiation_mean)) +
  geom_line(aes(colour = Series)) +
  geom_errorbar(aes(ymin = Radiation_mean - Radiation_sd,
                    ymax = Radiation_mean + Radiation_sd),
                width = 0.05,
                alpha = 0.3) +
  facet_grid(Series ~ Month,
             scales = "free_y") +
  labs(title = 'The Average Daily Irradiance',
       x = 'Time',
       y = 'Downwelling Irradiance (W/m^2)',
       subtitle = 'The mean daily irradiance per month',
       caption = 'includes standard deviation for each point') +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")  

```

Hopefully I've shown how easy it is to use r to transform our data and visualise it.

It wouldn't take much for me to create a template or a parametrised file to easily change inputs.

I also could easily feed this data into a solar model and use it to check for performance discrepancies due to weather.


I personally used Excel for all of this work before, and to be frank I can do it easily in that tool (I'll never betray Excel 😀), but my experience with Excel and r are vastly different. I've only truly started using r recently while I've used Excel from university and straight through my entire career.

This should hopefully hint at how useful r can be in an energy analyst's pipeline.

```{r Session Info}
sessionInfo() %>%
  print(., locale = FALSE)
```
